Real-Life Algebraic File Synchronization

### 2015-10-25 2015/paper.tex
Continued with the new algebra

\newcommand{\setv}{\mathcal{V}}
\newcommand{\setf}{\mathcal{V}_F}
\newcommand{\setd}{\mathcal{V}_D}
\newcommand{\setp}{\mathcal{P}}
\newcommand{\empt}{\bot}
\newcommand{\parent}{\mathtt{parent}}
\newcommand{\FS}{\mathrm{FS}}

@@

We model a filesystem using a function with a set of paths $\setp$ as its domain,
and a set of possible contents $\setv$ as its codomain:
\[ \FS : \setp \rightarrow \setv \] 
In our model, $\setp$ contains all possible paths, and $\setv$ contains a special
element, $\empt$, which is the value of $FS$ at paths where there are no files
or directories.
We consider the contents of files, as well as any meta-information on files
and directories (e.g. permissions or flags) part of the values in $\setv$.

@@

Moreover, in $\setv$ we distinguish between files ($\setf$) and directories ($\setd$), that is:
\[ \setv = \{\empt\} \cup \setf \cup \setd \]
The tree propety can then be defined as
\[ \forall \FS : \forall p\in\setp : \FS(p) \neq \empt \wedge \parent(p) \neq \empt \Rightarrow \FS(\parent(p)) \in \setd \]

% diagram

@@

\section{Commands and Sequences}

@@

Let us consider what kind of information may be encoded in the commands.
A usual set of commands, based on the most frequent tools implemented by filesystems,
might be $create(p,V)$, $edit(p,V)$ and $remove(p,V)$ where $p\in\setp$ and $V\in\setv$.
Clearly the commands need to contain information about the state they leave the filesystem
in at the path on which they operate, that is, they need to contain the ``end'' value, $V$.
Notice, however, that the commands above also encode some information about the filesystem
{\it before} the command is applied; namely, $create$ requires that there are no files
or directories at $p$, while $edit$ and $remove$ require the opposite.

@@

>>>>

\newcommand{\setv}{\mathcal{V}}
\newcommand{\setf}{\mathcal{V}_F}
\newcommand{\setd}{\mathcal{V}_D}
\newcommand{\setb}{\mathcal{V}_\empt}
\newcommand{\setp}{\mathcal{P}}
\newcommand{\empt}{\bot}
\newcommand{\parent}{\mathtt{parent}}
\newcommand{\FS}{\Phi} % {\mathrm{FS}}

\newcommand{\cbrk}{\mathtt{break}}
\newcommand{\cbb}{\overrightarrow{\empt\empt}}
\newcommand{\cbf}{\overrightarrow{\empt{F}}}
\newcommand{\cbd}{\overrightarrow{\empt{D}}}
\newcommand{\cfb}{\overrightarrow{F\empt}}
\newcommand{\cff}{\overrightarrow{FF}}
\newcommand{\cfd}{\overrightarrow{FD}}
\newcommand{\cdb}{\overrightarrow{D\empt}}
\newcommand{\cdf}{\overrightarrow{DF}}
\newcommand{\cdd}{\overrightarrow{DD}}
\newcommand{\cxy}{\overrightarrow{XY}}

@@

We model a filesystem using a function $\FS$ with a set of paths $\setp$ as its domain,
and a set of possible contents $\setv$ as its codomain:
\[ \FS : \setp \rightarrow \setv \] 
In our model, $\setp$ contains all possible paths, and $\setv$ contains a special
element, $\empt$, which is the value of $FS$ at paths where there are no files
or directories.
We consider the contents of files, as well as any meta-information on files
and directories (e.g. permissions or flags) part of the values in $\setv$.

@@

Moreover, in $\setv$ we distinguish between files ($\setf$) and directories ($\setd$), that is,
if $\setb = \{\empt\}$ then:
\[ \setv = \setb \cup \setf \cup \setd \]
The tree propety can then be defined as
\[ \forall \FS : \forall p\in\setp : \FS(p) \neq \empt \wedge \parent(p) \neq \empt \Rightarrow \FS(\parent(p)) \in \setd \]

% diagram
% /> D ---> F ---> BOT <\
% \--/               \--/

@@

\section{Commands}

@@

Let us consider what kind of information may be encoded in the commands.
A usual set of commands, based on the most frequent tools implemented by filesystems,
might be $create(p,V)$, $edit(p,V)$ and $remove(p)$ where $p\in\setp$ and $V\in\setv$.
Clearly the commands need to contain information about the state they leave the filesystem
in at the path on which they operate, that is, they need to contain the ``end'' value, $V$.
Notice, however, that the commands above also encode some information about the filesystem
{\it before} the command is applied; namely, $create$ requires that there are no files
or directories at $p$, while $edit$ and $remove$ require the opposite.

We also know that after $create$ or $edit$, $\FS(p)\neq\empt$, whereas after $remove$,
$\FS(p)$ will be $\empt$. However, from \cite{NREC:alg} we know that a useful set of axioms
will need to distinguish between edits that result in directories ($edit(p,D)$) and
ones that result in files ($edit(p,F)$). This creates a seemingly arbitrary asymmetry where
there is more information encoded into commands about their results than about the
original state of the filesystem.

In line with the aim to encode in commands as much information as possible,
and in order to resolve this asymmetry, we propose a set of commands that encode
the type of the original state of $\FS(p)$ as well.
(Some real-life filesystem commands like $mkdir$ or $rmdir$ do this already.)
Please note that there is never any need to encode information about the
{\it exact} value of path $p$ in a command, merely its type ($D$, $F$ or $\empt$),
as the success or failure of subsequent commands only depend on the type of the value.

We therefore propose to have a command for each pair of types.
For want of a better system, we will simply name our commands by concatenating
two of $D$, $F$ and $\empt$ with an arrow, where the left sign notes the type of value
in the filesystem before the command is applied, and the right sign notes the type
afterwards. For example, $mkdir(p,D)$ is $\cbd(p,D)$ and $rmdir(p)$ is $\cdb(p)$.

@@

\section{Applying Commands}

The commands can only succeed if the original value at $p$ has a type that matches
the type required by a command. If this is not the case, or if the resulting
filesystem no longer has the tree property, then we say that the command
\textbf{breaks} the filesystem. Broken filesystems are considered equal
(but not equal to any working filesystem), and we note them by $\FS=\empt$.

So that we could reason about sequences of commands that break every filesystem
in the algebra, we introduce the command $\cbrk$ that simply breaks every filesystem.
We note that a command or a sequence of commands $S$ is applied to a filesystem
by prefixing the command or sequence to it:
$S\FS$; or $\cbrk\FS=\empt$; or $\Xi = \cff(p,F)\FS$, for which $\Xi(p)=F$ holds
and $\Xi(q)=\FS(q)$ for any path $q\neq p$.

\section{Sequences of Commands}

\begin{thebibliography}{99}

\bibitem{NREC:alg} Ramsey, Norman and Elod Csirmaz: {\it An algebraic approach to
file synchronization...}

\end{thebibliography}
